---
title: "ggplot2"
author: "MianMian"
date: "2020/7/8"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(MASS)
library(conflicted)
conflict_prefer("select","dplyr")
library(readr)
#National Parks in California 
ca <- read_csv("data/ca.csv") 
#Acadia National Park 
acadia <- read_csv("data/acadia.csv") 
#Southeast US National Parks 
se <- read_csv("data/se.csv") 
#2016 Visitation for all Pacific West National Parks 
visit_16 <- read_csv("data/visit_16.csv") 
#All Nationally designated sites in Massachusetts 
mass <- read_csv("data/mass.csv")
```

To add a geom to the plot use + operator. Because we have two continuous variables, let’s use geom_point() first and then assign x and y aesthetics (aes):

```{r}
head(ca)
```

```{r}
ggplot(data = ca) + 
geom_point(aes(x = year, y = visitors,color=park_name))+
labs(x = "Year", y = "Visitation", title = "California National Park Visitation") + 
theme_linedraw()+
theme(legend.title=element_blank()) 
```

We can now make the faceted plot by splitting further by park using park_name (within a single plot):

```{r}
  ggplot(data = se) + geom_point(aes(x = year, y = visitors, color = park_name)) + 
  facet_wrap(~ state, scales = "free")+
  theme(legend.title=element_blank()) 
```

A geom is the geometrical object that a plot uses to represent data. People often describe plots by the type of geom that the plot uses. For example, bar charts use bar geoms, line charts use line geoms, boxplots use boxplot geoms, and so on. Scatterplots break the trend; they use the point geom. You can use different geoms to plot the same data. To change the geom in your plot, change the geom function that you add to ggplot(). Let’s look at a few ways of viewing the distribution of annual visitation (visitors) for each park (park_name).

```{r}
  ggplot(data = se) + 
  geom_jitter(aes(x = park_name, y = visitors, color = park_name), width = 0.1, alpha = 0.4) +          coord_flip() +theme(legend.position = "none")
```


```{r}
# xiang xian tu, heng xian wei zhong wei shu
ggplot(se, aes(x = park_name, y = visitors)) + geom_boxplot() + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

None of these are great for visualizing data over time. We can use
geom_line() in the same way we used geom_point.

```{r}
ggplot(se, aes(x = year, y = visitors, color = park_name)) + geom_line()
```


```{r}
ggplot(data = acadia) + geom_point(aes(x = year, y = visitors)) + 
  geom_line(aes(x = year, y = visitors),col='blue') + 
  geom_smooth(aes(x = year, y = visitors),se=F,col='red') + 
  labs(title = "Acadia National Park Visitation", y = "Visitation", x = "Year") + 
  theme_bw()
```
Next, let’s take a look at a bar chart. Bar charts seem simple, but they are interesting because they reveal something subtle about plots. Consider a basic bar chart, as drawn with geom_bar(). The following chart displays the total number of parks in each state within the Pacific West region.

```{r}
ggplot(data = visit_16, aes(x = state)) + geom_bar()
```

```{r}
ggplot(data = visit_16, aes(x = state, y = visitors, fill = park_name)) + 
  geom_bar(stat = "identity", position = "dodge")

```
After creating your plot, you can save it to a file in your favorite format. The Export tab in the Plot pane in RStudio will save your plots at low resolution, which will not be accepted by many journals and will not scale well for posters. Instead, use the ggsave() function, which allows you easily change the dimension and resolution of your plot by adjusting the appropriate arguments (width, height and dpi):

```{r}
my_plot <- ggplot(data = mass) + 
      geom_bar(aes(x = type, fill = park_name)) +
  labs(x = "", y = "")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 7))
ggsave("name_of_file.png", my_plot, width = 15, height = 10)

```

```{r}
pdf(file="name_of_file.pdf")
ggplot(data = mass) + 
      geom_bar(aes(x = type, fill = park_name)) +
  labs(x = "", y = "")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 7))

dev.off()
```